Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-09-02 bbbq_article.tex
Cite R packages I wrote

To determine the percentages of MHC-I and MHC-II epitopes overlap
with TMHs, we used mostly the same analysis as described in \cite{bianchi2017}.
To summarize: a proteome is converted into all its possible 9-mers. For each
of these peptides, it is measured if it has been part of a TMH (at least
1 residue), as well as if the peptide is a binder to an MHC-I haplotype.
Because the percentage of TMH residues is known, a binomial test can be
used to determine if TMH-derived peptides are as likely to be a binder
as expected by chance (the null hypothesis), or that TMHs are indeed an
overrepresented source of epitopes (as found by \cite{bianchi2017}).

@@

We compare the conservation values for TMH versus non-TMH locations,
using a Kolmogorov-Smirnov test, for a significance level of 5 percent.

@@

\begin{table}[]
  \begin{tabular}{lll}
    Goal & Tool  \\ 
    \hline
    Predict topology                  & TMHMM \cite{krogh2001predicting} \\
    Predict topology                  & PureseqTM \cite{wang2019efficient} \\
    Predict epitopes MHC-I            & \verb;epitope-prediction; \cite{bianchi2017} \\
    Predict epitopes MHC-II           & NetMHCIIpan \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
    Predict epitopes MHC-I and MHC-II & MHCnuggets \cite{shao2020high} \\
    Call TMHMM from R                 & tmhmm \\
    Call PureseqTM from R            & pureseqtmr \\
    Call NetMHCIIpan from R           & netmhc2pan \\
    Call MHCnuggets from R            & mhcnuggetsr \\
    Call TMHMM from R                 & tmhmm \\
    \verb;epitope-prediction; LUT     & epiprepreds \\
    NetMHCIIpan LUT                   & nmhc2ppreds \\
    MHCnuggets LUT                    & mhcnpreds \\
    Combine all                       & bbbq
  \end{tabular}
  \caption{
    Overview of all software used in this research.
    Software without a reference is written for this research.
    'LUT' is an abbreviation for 'Look-up table'. 
  }
  \label{table:software_used}
\end{table}

@@

For MHC-I, there are multiple computational tools developed 
to predict epitopes. 
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
A tool developed later is \verb;epitope-prediction; \cite{bianchi2017},
which uses a stabilized matrix method \cite{kim2009derivation}.
The most recent FOSS tool available is 
MHCnuggets \cite{shao2020high} (which can do both MHC-I 
and MHC-II prediction) which we will use in this study

@@

Also for MHC-II, there are multiple computational tools developed 
to predict epitopes,
such as using a trained neural network (\cite{nielsen2003reliable})
or a Gibbs sampling approach (\cite{nielsen2004improved}).
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
gave rise to the most accurate predictions.
Later tools are \cite{zhang2013predicting},
\cite{trolle2014nettepi}, \cite{zhang2015accurate}
and the very recent MHCnuggets (\cite{shao2020high}),
the latter which we will use in this study.

>>>>

To determine the percentages of MHC-I and MHC-II epitopes overlap
with TMHs, we used mostly the same analysis as described in \cite{bianchi2017}.
To summarize: from a proteome all possible 9-mers are derived. For each
of these peptides, it is measured if it has been part of a 
TMH (at least overlapping 1 residue), 
as well as if the peptide is a binder to an MHC-I haplotype.
Because the percentage of TMH residues is known, a binomial test can be
used to determine if TMH-derived peptides are as likely to be a binder
as expected by chance (the null hypothesis), or that TMHs are indeed an
overrepresented source of epitopes (as found by \cite{bianchi2017}).

@@

We compare the conservation values for TMH versus non-TMH locations,
for all membrane proteins combined,
using a Kolmogorov-Smirnov test, for a significance level of 5 percent.

@@

\begin{table}[]
  \begin{tabular}{lll}
    Goal & Tool  \\ 
    \hline
    Predict topology                  & TMHMM \cite{krogh2001predicting} \\
    Predict topology                  & PureseqTM \cite{wang2019efficient} \\
    Predict epitopes MHC-I            & \verb;epitope-prediction; \cite{bianchi2017} \\
    Predict epitopes MHC-II           & NetMHCIIpan \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
    Predict epitopes MHC-I and MHC-II & MHCnuggets \cite{shao2020high} \\
    Call TMHMM from R                 & tmhmm \cite{tmhmm} \\
    Call PureseqTM from R             & pureseqtmr \cite{pureseqtmr} \\
    Call NetMHCIIpan from R           & netmhc2pan \cite{netmhc2pan} \\
    Call MHCnuggets from R            & mhcnuggetsr \cite{mhcnuggetsr} \\
    \verb;epitope-prediction; LUT     & epiprepreds \cite{epiprepreds} \\
    NetMHCIIpan LUT                   & nmhc2ppreds \cite{nmhc2ppreds} \\
    MHCnuggets LUT                    & mhcnpreds \cite{mhcnpreds} \\
    Combine all                       & bbbq \cite{bbbq}
  \end{tabular}
  \caption{
    Overview of all software used in this research.
    Software without a reference is written for this research.
    'LUT' is an abbreviation for 'Look-up table'. 
  }
  \label{table:software_used}
\end{table}

@@

For MHC-I, there are multiple computational tools developed 
to predict epitopes. 
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
A tool developed later is \verb;epitope-prediction; \cite{bianchi2017},
which uses a stabilized matrix method \cite{kim2009derivation}.
The most recent FOSS tool available is 
MHCnuggets \cite{shao2020high} (which can do both MHC-I 
and MHC-II prediction) which is used in this study

@@

Also for MHC-II, there are multiple computational tools developed 
to predict epitopes,
such as using a trained neural network (\cite{nielsen2003reliable})
or a Gibbs sampling approach (\cite{nielsen2004improved}).
According to \cite{lundegaard2011prediction}, in 2011,
from a set of multiple tools, 
NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
gave rise to the most accurate predictions.
Later tools are \cite{zhang2013predicting},
\cite{trolle2014nettepi}, \cite{zhang2015accurate}
and the very recent MHCnuggets (\cite{shao2020high}),
the latter which is used in this study.
